---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)

steps <- read.csv("activity.csv") %>% mutate(date=as.Date(date))
summary(steps)
```

## What is mean total number of steps taken per day?

```{r}


stepsperday <- steps %>% group_by(date) %>% summarise(steps=sum(steps))
meanstepperday <- mean(stepsperday$steps, na.rm = T)
medianstepperday <- median(stepsperday$steps, na.rm = T)

ggplot(stepsperday, aes(x=steps)) + geom_histogram(binwidth=500)

```

The mean total number of steps taken per day is equal `r meanstepperday` and the median number of steps is `r medianstepperday`.

## What is the average daily activity pattern?

```{r}
avgstepsperint <- steps %>% group_by(interval) %>% summarise(steps=mean(steps, na.rm = T))
ggplot(avgstepsperint, aes(x = interval, y = steps)) + geom_line()


maxstepint <- last(avgstepsperint$interval, order_by = avgstepsperint$steps)
maxstep <- max(avgstepsperint$steps)
```

The `r maxstepint` is the interval with the highest average number of steps (`r maxstep`).

## Imputing missing values

```{r}
narows <- nrow(steps) - sum(complete.cases(steps))
```
Dataset contains `r narows` rows with missing data. I will replace NAs with intervals mean rounded to the closest integer.

```{r}
nonasteps <- steps %>% inner_join(avgstepsperint, by='interval') %>% mutate(steps.z=round(steps.y), steps.z=as.integer(steps.z), steps=coalesce(steps.x, steps.z)) %>% select(steps, date, interval)

nonastepsperday <- nonasteps %>% group_by(date) %>% summarise(steps=sum(steps))
nonameanstepperday <- mean(nonastepsperday$steps)
nonamedianstepperday <- median(nonastepsperday$steps)

ggplot(nonastepsperday, aes(x=steps)) + geom_histogram(binwidth=500)

```

After filling NAs we can see a new 'spike' around mean (10.7K). The mean and median stayed almost unafected. The new mean total number of steps taken per day is equal `r nonameanstepperday` and the new median number of steps is `r nonamedianstepperday`.


## Are there differences in activity patterns between weekdays and weekends?

```{r} 
nonasteps$daytype <- as.factor(ifelse(wday(nonasteps$date) %in% c(1,7), "weekend", "weekday"))
nonaavgstepsperint <- nonasteps %>% group_by(daytype, interval) %>% summarise(steps=mean(steps))
ggplot(nonaavgstepsperint, aes(x = interval, y = steps)) + geom_line() + facet_wrap( ~ daytype)
```

During the weekday there is a spike in the beginig of the day suggesting people exercise before or while comuting to work. During the weekend steps are distributed more evenly all day long.

